### 2020-07-05 chapters/validation.tex
Multiple grammar corrections

All the synthetic datasets used throughout this section are a time-series where at each time-step we sample uniformly from a fixed distribution, \textit{i.e.} there are no time correlations in the used datasets. The reference dataset R1 used in this first experiment contained a single feature that followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$ for the 5 million rows of the dataset. Figure \ref{fig:timeseries-r1} shows the time-series of the single feature in dataset R1. The dataset processed in the streaming phase (\textit{i.e.} the target dataset) was generated with a size of 5 million events and a single feature. For the first half, \textit{i.e.}, for the first 2.5 million events, the feature followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$, like the reference dataset. For the other half, we changed the generating distribution to be a continuous uniform one with lower bound $a=100$ and upper bound $b=200$. Figure \ref{fig:timeseries-t1} shows the time-series of the single feature in the target dataset T1 where we see the abrupt change in feature values.
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/01-reference.pdf}
  \captionof{figure}{Time-series for dataset R1}
  \label{fig:timeseries-r1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/01-target.pdf}
  \captionof{figure}{Time-series for dataset T1}
  \label{fig:timeseries-t1}
\end{minipage}
\end{center}

@@

In this experiment, we used a half-life $n_{1/2}=62500$ and corresponding sample size of 250,000 events, like in the previous Section' experiment 02. Figure \ref{fig:JSD-signal-test02} shows the computed JSD signal for the target dataset and Figure \ref{fig:JSD-signal-zoom-test02} zooms in on the first 2 millionth events, before any distribution is changed. We observe that once again we have no false positives in the first 2 million events, which is ideal. 
\begin{figure}[!htb]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/stream-analysis-viz-test02.pdf}
  \caption{For the whole time-series}
  \label{fig:JSD-signal-test02}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/stream-analysis-viz-zoom-test02.pdf}
  \caption{Zoom in on the first 2 million events}
  \label{fig:JSD-signal-zoom-test02}
\end{subfigure}
\caption{JSD signal and threshold}
\end{figure}
Furthermore, we observe that the computed JSD signal varies accordingly with the changes in distribution. Despite us changing the distributions to subsets of the previous ones to try and deceive our method, our signal accurately represents these changes. In Figure \ref{fig:JSD-signal-test02}, we see several growth points that correspond to the points where the generating gaussian distribution parameters changed. Hence, we are able to detect changes in distribution parameter changes and quantify how big the change is (the closer to 1 our signal is).

@@

We used the reference dataset R5 (Figure \ref{fig:timeseries-r5}) which had 5 million events, with a single-feature that followed a gaussian with  $\mu=100$ and $\sigma=10$. Our target dataset T5 is represented in Figure \ref{fig:timeseries-t5}. For the first 2 million events, the generating distribution used was similar to the one used in the reference dataset, a gaussian with $\mu=100$ and $\sigma=10$. However, at the 2 millionth event, we change the gaussian standard deviation to $\sigma=30$. Then, at the 4 millionth event, we change the distribution type. We use an uniform distribution with lower bound $a=80$ and upper bound $b=120$. Note in Figure \ref{fig:timeseries-t5} that this last portion resembles the first one, the reference.
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/timeseries-r5.pdf}
  \captionof{figure}{Time-series for dataset R5}
  \label{fig:timeseries-r5}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/timeseries-t5.pdf}
  \captionof{figure}{Time-series for dataset T5}
  \label{fig:timeseries-t5}
\end{minipage}
\end{center}

@@

\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|ll}
        \cline{1-3}
        \textbf{Feature} & \textbf{Reference Distribution} & \textbf{Parameters} &  &  \\ \cline{1-3}
        x1               & Gaussian/Normal                 & $\mu=100$, $\sigma=5$  &  &  \\ \cline{1-3}
        x2               & Gaussian/Normal                 & $\mu=100$, $\sigma=5$  &  &  \\ \cline{1-3}
        x3               & Uniform                         & $a=200$, $b=250$       &  &  \\ \cline{1-3}
        x4               & Uniform                         & $a=50$, $b=125$        &  &  \\ \cline{1-3}
        \end{tabular}
    \end{center}
    \caption{Multi feature reference dataset and feature distributions}
    \label{tbl:multi-feat-ref-dataset-distros}
\end{table}

\begin{figure}[!htb] 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x1.pdf} 
    \caption{Feature x1 reference time-series} 
    \label{fig:timeseries-r6-x1} 
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x2.pdf} 
    \caption{Feature x2 reference time-series} 
    \label{fig:timeseries-r6-x2} 
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x3.pdf} 
    \caption{Feature x3 reference time-series} 
    \label{fig:timeseries-r6-x3} 
    \vspace{4ex}
  \end{minipage}%% 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x4.pdf} 
    \caption{Feature x4 reference time-series} 
    \label{fig:timeseries-r6-x4} 
    \vspace{4ex}
  \end{minipage} 
\end{figure}

We used a target dataset T6 with 5 million events and the same four features ($x_1$, $x_2$, $x_3$ and $x_4$). Feature x1 followed the reference distribution for the 5 million events of the target dataset. We did not change this feature underlying distribution to serve as a control, \textit{i.e.}, we expect no alerts for this one. Figure \ref{fig:timeseries-t6-x1} shows the time-series for the target dataset T6 and feature $x_1$.

Feature x2's underlying distribution changed four times and returned to normal, so we expect to see alarms with increasing divergence values but eventually no more alarms. Figure \ref{fig:timeseries-t6-x2} shows the time-series for the target dataset T6 and feature $x_2$. Feature $x_2$ followed a gaussian with $\mu=100$ and $\sigma=5$ like in the reference period up to the 1 millionth event mark, then a gaussian with $\mu=100$ and $\sigma=10$ up to the 2.5 millionth event, then a gaussian $\mu=100$ and $\sigma=15$ until the 3 millionth event, then a gaussian $\mu=100$ and $\sigma=20$ until the 4 millionth event and finally returned to the original gaussian with $\mu=100$ and $\sigma=5$ until the last event. Table \ref{tbl:multi-feat-x2-changes} sums up the distributions used and its parameters for each interval of the reference dataset for feature x2.
\begin{table}[!htb]
    \begin{center}
    \begin{tabular}{|c|c|c|c|l}
    \cline{1-4}
    \textbf{From Event} & \textbf{To Event} & \textbf{Distribution} & \multicolumn{1}{l|}{\textbf{Parameters}} &  \\ \cline{1-4}
    0                   & 1,000,000         & Gaussian              & $\mu=100, \sigma=5$                      &  \\ \cline{1-4}
    1,000,000           & 2,500,000         & Gaussian              & $\mu=100, \sigma=10$                     &  \\ \cline{1-4}
    2,500,000           & 3,000,000         & Gaussian              & $\mu=100, \sigma=15$                     &  \\ \cline{1-4}
    3,000,000           & 4,000,000         & Gaussian              & $\mu=100, \sigma=20$                      &  \\ \cline{1-4}
    4,000,000           & 5,000,000         & Gaussian              & $\mu=100, \sigma=5$                      &  \\ \cline{1-4}
    \end{tabular}
    \end{center}
    \caption{Feature x2 generating distributions}
    \label{tbl:multi-feat-x2-changes}
\end{table}

Feature x3's underlying distribution changed three times and never returned to normal, so we expect to see alarms with increasing divergence values up until the end of the stream analysis. Figure \ref{fig:timeseries-t6-x3} shows the time-series for the target dataset T6 and feature x3. Feature x3 followed a uniform distribution with $a=200$ and $b=250$ like in the reference period up to the 1.5 millionth event mark, then a uniform distribution with $a=200$ and $b=220$ up to the 2 millionth event, then a uniform distribution with $a=180$ and $b=200$ until the 3.5 millionth event and finally a gaussian with $\mu=50$ and $\sigma=20$ until the last event. Table \ref{tbl:multi-feat-x3-changes} sums up the distributions used and its parameters for each interval of the reference dataset for feature x3.
\begin{table}[!htb]
    \begin{center}
    \begin{tabular}{|c|c|c|c|l}
    \cline{1-4}
    \textbf{From Event} & \textbf{To Event} & \textbf{Distribution} & \multicolumn{1}{l|}{\textbf{Parameters}} &  \\ \cline{1-4}
    0                   & 1,500,000         & Uniform               & $a=200, b=250$                           &  \\ \cline{1-4}
    1,500,000           & 2,000,000         & Uniform               & $a=200, b=220$                           &  \\ \cline{1-4}
    2,000,000            & 3,500,000         & Uniform               & $a=180, b=200$                           &  \\ \cline{1-4}
    3,500,000           & 5,000,000         & Gaussian              & $\mu=50, \sigma=20$                      &  \\ \cline{1-4}
    \end{tabular}
    \end{center}
    \caption{Feature x3 generating distributions}
    \label{tbl:multi-feat-x3-changes}
\end{table}

Feature x4's underlying distribution changed two times and returned to normal, so we expect to see no alarms towards the end of the stream analysis. Figure \ref{fig:timeseries-t6-x4} shows the time-series for the target dataset T6 and feature x4. Feature x4 followed a uniform distribution like the reference one, with $a=50$ and $b=125$ up until the 1.5 millionth event. After that, it followed a uniform distribution with $a=125$ and $b=200$ until the 3 millionth event where it resumed its original uniform distribution (with $a=50$ and $b=125$) until the end. Table \ref{tbl:multi-feat-x4-changes} sums up the distributions used and its parameters for each interval of the reference dataset for feature x4.
\begin{table}[!htb]
    \begin{center}
    \begin{tabular}{|c|c|c|c|l}
    \cline{1-4}
    \textbf{From Event} & \textbf{To Event} & \textbf{Distribution} & \multicolumn{1}{l|}{\textbf{Parameters}} &  \\ \cline{1-4}
    0                   & 1,500,000         & Uniform               & $a=50, b=125$                            &  \\ \cline{1-4}
    1,500,000           & 3,000,000         & Uniform               & $a=125, b=200$                           &  \\ \cline{1-4}
    3,000,000           & 5,000,000         & Uniform               & $a=50, b=125$                            &  \\ \cline{1-4}
    \end{tabular}
    \end{center}
    \caption{Feature x4 generating distributions}
    \label{tbl:multi-feat-x4-changes}
\end{table}

\begin{figure}[!htb] 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x1.pdf} 
    \caption{Feature x1 target time-series} 
    \label{fig:timeseries-t6-x1} 
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x2.pdf} 
    \caption{Feature x2 target time-series} 
    \label{fig:timeseries-t6-x2} 
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x3.pdf} 
    \caption{Feature x3 target time-series} 
    \label{fig:timeseries-t6-x3} 
    \vspace{4ex}
  \end{minipage}%% 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x4.pdf} 
    \caption{Feature x4 target time-series} 
    \label{fig:timeseries-t6-x4} 
    \vspace{4ex}
  \end{minipage} 
\end{figure}

In this experiment, we used a tuple-based half-life $n_{1/2}=62500$, \textit{i.e.} sampling and target windows of 250,000 tuples. We used 100 bins for the reference histogram plus the two special infinity bins. In this multi-feature scenario, we used the Holm-Bonferroni correction discussed in Section \ref{sec:holmbonferroni} and set the Family-Wise Error Rate (FWER) to 1\% (FWER = 0.01). Furthermore, instead of doing 1000 samples as we did previously in an ad-hoc way, we use Equation \ref{eq:optimal-n-samples} to determine the minimum number of samples to make while ensuring $\gamma=1\%$ (recall Section \ref{sec:nsamples}). For our four feature analysis scenario, using as inputs  $\gamma=0.01$ and $FWER=0.01$ in Equation \ref{eq:optimal-n-samples} tells us we need to make 1840 samples minimum. Using 300 Spark executors, each given 3G of memory, we finished the batch phase in 418 seconds.

In Section \ref{sec:stream-phase}, we mention that, in our stream analysis, we perform our multiple hypothesis divergence tests on a fixed user-defined frequency. Finding an optimal frequency value for the test falls out of scope for this Thesis, but ideally, we set it to be large enough so that we do fewer computations and achieve higher throughput but still get alerts in useful time. In this experiment, we performed the divergence test with the Holm-Bonferroni correction \ref{sec:holmbonferroni} for every new event (frequency of 1 event). We did this to collect a list of ordered corrected p-values for each feature. A corrected p-value (derived from Equation \ref{eq:corrected-pvalue}) is given by:
\begin{equation}
    p-value_{corrected} = p-value_{distance_{x}} * (m - k)
    \label{for:corrected-pvalue}
\end{equation}
With this list, for each feature, we can plot the corrected p-values on each event and compare it with the FWER threshold $\alpha$. Note that despite performing the test at the maximum frequency (in real-time) we still achieved a throughput of 17,921 transactions per second (TPS), even if only for four features.

@@

In Figure \ref{fig:x1-corrected-pvalues} we see the plot of corrected p-values for feature x1 and the probability threshold black horizontal line at the bottom. Recall that feature x1' values were generated by a Gaussian distribution for both the reference period (Figure \ref{fig:timeseries-r6-x1}) and the target period (Figure \ref{fig:timeseries-t6-x1}), with the same parameters. We observe that the corrected p-values never pass below the threshold, as we expected. Hence, this feature is never alerted as divergent.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x1-corrected-pvalues.png}
      \caption{Feature x1 corrected p-values and threshold}
      \label{fig:x1-corrected-pvalues}
    \end{center}
\end{figure}

Figure \ref{fig:x2-corrected-pvalues} shows a plot of feature x2's corrected p-values and the FWER threshold. Feature x2 underlying distribution was changed four times (Table \ref{tbl:multi-feat-x2-changes}): at the 1, 2.5, 3 and 4 millionth events. The first change is consistent with the sudden drop in corrected p-value for this feature, as we cross below the threshold and enter a divergent state. Around a thousand events after the change we saw p-values of 0.2\% (or 0.002). The values keep dropping as time passes and as we introduce bigger changes in distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x2-corrected-pvalues.png}
      \caption{Feature x2 corrected p-values and threshold}
      \label{fig:x2-corrected-pvalues}
    \end{center}
\end{figure}
We resume the reference or original distribution at the 4 millionth event. However, as we can see from a close-up on the original plot (Figure \ref{fig:x2-corrected-pvalues-zoom}), it takes around 600,000 events for the target sliding histogram aggregation to once again resemble the reference one, resulting in a high enough p-value and thus stop alerting x2 as divergent.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x2-corrected-pvalues-zoom2.png}
      \caption{Feature x2 zoomed in corrected p-values and threshold}
      \label{fig:x2-corrected-pvalues-zoom}
    \end{center}
\end{figure}

Feature x3 was the only feature that never returned to its original distribution and is the only one that should be reported until the end as divergent. This is in fact the case as illustrated in Figure \ref{fig:x3-corrected-pvalues}. The first change to x3's generating distribution was introduced at the 1.5 millionth event and we clearly see the corrected p-value rapidly approaching values close to zero. For the rest of the dataset, this x3's p-value remains below the threshold, as we never swap to its original distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x3-corrected-pvalues.png}
      \caption{Feature x3 corrected p-values and threshold}
      \label{fig:x3-corrected-pvalues}
    \end{center}
\end{figure}

>>>>

All the synthetic datasets used throughout this section are a time-series where at each time-step we sample uniformly from a fixed distribution, \textit{i.e.}, there are no time correlations in the used datasets. The reference dataset R1 used in this first experiment contained a single feature that followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$ for the 5 million rows of the dataset. Figure \ref{fig:timeseries-r1} shows the time-series of the single feature in dataset R1. The dataset processed in the streaming phase (\textit{i.e.}, the target dataset) was generated with a size of 5 million events and a single feature. For the first half, \textit{i.e.}, for the first 2.5 million events, the feature followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$, like the reference dataset. For the other half, we changed the generating distribution to be a continuous uniform one with lower bound $a=100$ and upper bound $b=200$. Figure \ref{fig:timeseries-t1} shows the time-series of the single feature in the target dataset T1 where we see the abrupt change in feature values.
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/01-reference.pdf}
  \captionof{figure}{Time-series for dataset R1}
  \label{fig:timeseries-r1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/01-target.pdf}
  \captionof{figure}{Time-series for dataset T1}
  \label{fig:timeseries-t1}
\end{minipage}
\end{center}

@@

In this experiment, we used a half-life $n_{1/2}=62500$ and corresponding sample size of 250,000 events, like in the previous section's experiment 02. Figure \ref{fig:JSD-signal-test02} shows the computed JSD signal for the target dataset and Figure \ref{fig:JSD-signal-zoom-test02} zooms in on the first 2 millionth events, before any distribution is changed. We observe that once again we have no false positives in the first 2 million events, which is ideal. 
\begin{figure}[!htb]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/stream-analysis-viz-test02.pdf}
  \caption{For the whole time-series}
  \label{fig:JSD-signal-test02}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/stream-analysis-viz-zoom-test02.pdf}
  \caption{Zoom in on the first 2 million events}
  \label{fig:JSD-signal-zoom-test02}
\end{subfigure}
\caption{JSD signal and threshold}
\end{figure}
Furthermore, we observe that the computed JSD signal varies accordingly with the changes in distribution. Despite us changing the distributions to subsets of the previous ones to try and deceive our method, our signal accurately represents these changes. In Figure \ref{fig:JSD-signal-test02}, we see several growth points that correspond to the points where the generating gaussian distribution parameters changed. Hence, we are able to detect changes in distribution parameter changes and quantify how big the change is (the closer to 1 our signal is).

@@

We used the reference dataset R5 (Figure \ref{fig:timeseries-r5}) which had 5 million events, with a single feature that followed a gaussian with  $\mu=100$ and $\sigma=10$. Our target dataset T5 is represented in Figure \ref{fig:timeseries-t5}. For the first 2 million events, the generating distribution used was similar to the one used in the reference dataset, a gaussian with $\mu=100$ and $\sigma=10$. However, at the 2 millionth event, we change the gaussian standard deviation to $\sigma=30$. Then, at the 4 millionth event, we change the distribution type. We use an uniform distribution with lower bound $a=80$ and upper bound $b=120$. Note in Figure \ref{fig:timeseries-t5} that this last portion resembles the first one, the reference.
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/timeseries-r5.pdf}
  \captionof{figure}{Time-series for dataset R5}
  \label{fig:timeseries-r5}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/timeseries-t5.pdf}
  \captionof{figure}{Time-series for dataset T5}
  \label{fig:timeseries-t5}
\end{minipage}
\end{center}

@@

\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|ll}
        \cline{1-3}
        \textbf{Feature} & \textbf{Reference Distribution} & \textbf{Parameters} &  &  \\ \cline{1-3}
        $x_1$               & Gaussian/Normal                 & $\mu=100$, $\sigma=5$  &  &  \\ \cline{1-3}
        $x_2$               & Gaussian/Normal                 & $\mu=100$, $\sigma=5$  &  &  \\ \cline{1-3}
        $x_3$               & Uniform                         & $a=200$, $b=250$       &  &  \\ \cline{1-3}
        $x_4$               & Uniform                         & $a=50$, $b=125$        &  &  \\ \cline{1-3}
        \end{tabular}
    \end{center}
    \caption{Multi feature reference dataset and feature distributions}
    \label{tbl:multi-feat-ref-dataset-distros}
\end{table}

\begin{figure}[!htb] 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x1.pdf} 
    \caption{Feature $x_1$ reference time-series} 
    \label{fig:timeseries-r6-x1} 
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x2.pdf} 
    \caption{Feature $x_2$ reference time-series} 
    \label{fig:timeseries-r6-x2} 
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x3.pdf} 
    \caption{Feature $x_3$ reference time-series} 
    \label{fig:timeseries-r6-x3} 
    \vspace{4ex}
  \end{minipage}%% 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-r6-x4.pdf} 
    \caption{Feature $x_4$ reference time-series} 
    \label{fig:timeseries-r6-x4} 
    \vspace{4ex}
  \end{minipage} 
\end{figure}

We used a target dataset T6 with 5 million events and the same four features ($x_1$, $x_2$, $x_3$ and $x_4$). Feature $x_1$ followed the reference distribution for the 5 million events of the target dataset. We did not change this feature underlying distribution to serve as a control, \textit{i.e.}, we expect no alerts for this one. Figure \ref{fig:timeseries-t6-x1} shows the time-series for the target dataset T6 and feature $x_1$.

Feature $x_2$'s underlying distribution changed four times and returned to normal, so we expect to see alarms with increasing divergence values but eventually no more alarms. Figure \ref{fig:timeseries-t6-x2} shows the time-series for the target dataset T6 and feature $x_2$. Feature $x_2$ followed a gaussian with $\mu=100$ and $\sigma=5$ like in the reference period up to the 1 millionth event mark, then a gaussian with $\mu=100$ and $\sigma=10$ up to the 2.5 millionth event, then a gaussian $\mu=100$ and $\sigma=15$ until the 3 millionth event, then a gaussian $\mu=100$ and $\sigma=20$ until the 4 millionth event and finally returned to the original gaussian with $\mu=100$ and $\sigma=5$ until the last event. Table \ref{tbl:multi-feat-x2-changes} sums up the distributions used and its parameters for each interval of the reference dataset for feature $x_2$.
\begin{table}[!htb]
    \begin{center}
    \begin{tabular}{|c|c|c|c|l}
    \cline{1-4}
    \textbf{From Event} & \textbf{To Event} & \textbf{Distribution} & \multicolumn{1}{l|}{\textbf{Parameters}} &  \\ \cline{1-4}
    0                   & 1,000,000         & Gaussian              & $\mu=100, \sigma=5$                      &  \\ \cline{1-4}
    1,000,000           & 2,500,000         & Gaussian              & $\mu=100, \sigma=10$                     &  \\ \cline{1-4}
    2,500,000           & 3,000,000         & Gaussian              & $\mu=100, \sigma=15$                     &  \\ \cline{1-4}
    3,000,000           & 4,000,000         & Gaussian              & $\mu=100, \sigma=20$                      &  \\ \cline{1-4}
    4,000,000           & 5,000,000         & Gaussian              & $\mu=100, \sigma=5$                      &  \\ \cline{1-4}
    \end{tabular}
    \end{center}
    \caption{Feature $x_2$ generating distributions}
    \label{tbl:multi-feat-x2-changes}
\end{table}

Feature $x_3$'s underlying distribution changed three times and never returned to normal, so we expect to see alarms with increasing divergence values up until the end of the stream analysis. Figure \ref{fig:timeseries-t6-x3} shows the time-series for the target dataset T6 and feature $x_3$. Feature $x_3$ followed a uniform distribution with $a=200$ and $b=250$ like in the reference period up to the 1.5 millionth event mark, then a uniform distribution with $a=200$ and $b=220$ up to the 2 millionth event, then a uniform distribution with $a=180$ and $b=200$ until the 3.5 millionth event and finally a gaussian with $\mu=50$ and $\sigma=20$ until the last event. Table \ref{tbl:multi-feat-x3-changes} sums up the distributions used and its parameters for each interval of the reference dataset for feature $x_3$.
\begin{table}[!htb]
    \begin{center}
    \begin{tabular}{|c|c|c|c|l}
    \cline{1-4}
    \textbf{From Event} & \textbf{To Event} & \textbf{Distribution} & \multicolumn{1}{l|}{\textbf{Parameters}} &  \\ \cline{1-4}
    0                   & 1,500,000         & Uniform               & $a=200, b=250$                           &  \\ \cline{1-4}
    1,500,000           & 2,000,000         & Uniform               & $a=200, b=220$                           &  \\ \cline{1-4}
    2,000,000            & 3,500,000         & Uniform               & $a=180, b=200$                           &  \\ \cline{1-4}
    3,500,000           & 5,000,000         & Gaussian              & $\mu=50, \sigma=20$                      &  \\ \cline{1-4}
    \end{tabular}
    \end{center}
    \caption{Feature $x_3$ generating distributions}
    \label{tbl:multi-feat-x3-changes}
\end{table}

Feature $x_4$'s underlying distribution changed two times and returned to normal, so we expect to see no alarms towards the end of the stream analysis. Figure \ref{fig:timeseries-t6-x4} shows the time-series for the target dataset T6 and feature $x_4$. Feature $x_4$ followed a uniform distribution like the reference one, with $a=50$ and $b=125$ up until the 1.5 millionth event. After that, it followed a uniform distribution with $a=125$ and $b=200$ until the 3 millionth event where it resumed its original uniform distribution (with $a=50$ and $b=125$) until the end. Table \ref{tbl:multi-feat-x4-changes} sums up the distributions used and its parameters for each interval of the reference dataset for feature $x_4$.
\begin{table}[!htb]
    \begin{center}
    \begin{tabular}{|c|c|c|c|l}
    \cline{1-4}
    \textbf{From Event} & \textbf{To Event} & \textbf{Distribution} & \multicolumn{1}{l|}{\textbf{Parameters}} &  \\ \cline{1-4}
    0                   & 1,500,000         & Uniform               & $a=50, b=125$                            &  \\ \cline{1-4}
    1,500,000           & 3,000,000         & Uniform               & $a=125, b=200$                           &  \\ \cline{1-4}
    3,000,000           & 5,000,000         & Uniform               & $a=50, b=125$                            &  \\ \cline{1-4}
    \end{tabular}
    \end{center}
    \caption{Feature $x_4$ generating distributions}
    \label{tbl:multi-feat-x4-changes}
\end{table}

\begin{figure}[!htb] 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x1.pdf} 
    \caption{Feature $x_1$ target time-series} 
    \label{fig:timeseries-t6-x1} 
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x2.pdf} 
    \caption{Feature $x_2$ target time-series} 
    \label{fig:timeseries-t6-x2} 
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x3.pdf} 
    \caption{Feature $x_3$ target time-series} 
    \label{fig:timeseries-t6-x3} 
    \vspace{4ex}
  \end{minipage}%% 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=1\linewidth]{figures/timeseries-t6-x4.pdf} 
    \caption{Feature $x_4$ target time-series} 
    \label{fig:timeseries-t6-x4} 
    \vspace{4ex}
  \end{minipage} 
\end{figure}

In this experiment, we used a tuple-based half-life $n_{1/2}=62500$, \textit{i.e.}, sampling and target windows of 250,000 tuples. We used 100 bins for the reference histogram plus the two special infinity bins. In this multi-feature scenario, we used the Holm-Bonferroni correction discussed in Section \ref{sec:holmbonferroni} and set the Family-Wise Error Rate (FWER) to 1\% (FWER = 0.01). Furthermore, instead of doing 1000 samples as we did previously in an ad-hoc way, we use Equation \ref{eq:optimal-n-samples} to determine the minimum number of samples to make while ensuring $\gamma=1\%$ (recall Section \ref{sec:nsamples}). For our four feature analysis scenario, using as inputs  $\gamma=0.01$ and $FWER=0.01$ in Equation \ref{eq:optimal-n-samples} tells us we need to make 1840 samples minimum. Using 300 Spark executors, each given 3G of memory, we finished the batch phase in 418 seconds.

In Section \ref{sec:stream-phase}, we mention that, in our stream analysis, we perform our multiple hypothesis divergence tests on a fixed user-defined frequency. Finding an optimal frequency value for the test falls out of scope for this thesis, but ideally, we set it to be large enough so that we do fewer computations and achieve higher throughput but still get alerts in useful time. In this experiment, we performed the divergence test with the Holm-Bonferroni correction \ref{sec:holmbonferroni} for every new event (frequency of 1 event). We did this to collect a list of ordered corrected p-values for each feature. A corrected p-value (derived from Equation \ref{eq:corrected-pvalue}) is given by:
\begin{equation}
    p-value_{corrected} = p-value_{distance_{x}} * (m - k)
    \label{for:corrected-pvalue}
\end{equation}
With this list, for each feature, we can plot the corrected p-values on each event and compare it with the FWER threshold $\alpha$. Note that despite performing the test at the maximum frequency (in real-time) we still achieved a throughput of 17,921 transactions per second (TPS), even if only for four features.

@@

In Figure \ref{fig:x1-corrected-pvalues} we see the plot of corrected p-values for feature $x_1$ and the probability threshold black horizontal line at the bottom. Recall that feature $x_1$'s values were generated by a Gaussian distribution for both the reference period (Figure \ref{fig:timeseries-r6-x1}) and the target period (Figure \ref{fig:timeseries-t6-x1}), with the same parameters. We observe that the corrected p-values never pass below the threshold, as we expected. Hence, this feature is never alerted as divergent.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x1-corrected-pvalues.png}
      \caption{Feature $x_1$ corrected p-values and threshold}
      \label{fig:x1-corrected-pvalues}
    \end{center}
\end{figure}

Figure \ref{fig:x2-corrected-pvalues} shows a plot of feature $x_2$'s corrected p-values and the FWER threshold. Feature $x_2$ underlying distribution was changed four times (Table \ref{tbl:multi-feat-x2-changes}): at the 1, 2.5, 3 and 4 millionth events. The first change is consistent with the sudden drop in corrected p-value for this feature, as we cross below the threshold and enter a divergent state. Around a thousand events after the change we saw p-values of 0.2\% (or 0.002). The values keep dropping as time passes and as we introduce bigger changes in distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x2-corrected-pvalues.png}
      \caption{Feature $x_2$ corrected p-values and threshold}
      \label{fig:x2-corrected-pvalues}
    \end{center}
\end{figure}
We resume the reference or original distribution at the 4 millionth event. However, as we can see from a close-up on the original plot (Figure \ref{fig:x2-corrected-pvalues-zoom}), it takes around 600,000 events for the target sliding histogram aggregation to once again resemble the reference one, resulting in a high enough p-value and thus stop alerting $x_2$ as divergent.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x2-corrected-pvalues-zoom2.png}
      \caption{Feature $x_2$ zoomed in corrected p-values and threshold}
      \label{fig:x2-corrected-pvalues-zoom}
    \end{center}
\end{figure}

Feature $x_3$ was the only feature that never returned to its original distribution and is the only one that should be reported until the end as divergent. This is in fact the case as illustrated in Figure \ref{fig:x3-corrected-pvalues}. The first change to $x_3$'s generating distribution was introduced at the 1.5 millionth event and we clearly see the corrected p-value rapidly approaching values close to zero. For the rest of the dataset, this $x_3$'s p-value remains below the threshold, as we never swap to its original distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.8]{figures/x3-corrected-pvalues.png}
      \caption{Feature $x_3$ corrected p-values and threshold}
      \label{fig:x3-corrected-pvalues}
    \end{center}
\end{figure}
